---
title: "Pre-process MSUPRP sample data for BATools"
output: html_notebook
---

# Information about the dataset

The data comes from https://github.com/steibelj/gwaR, an example dataset for gwas study for MSUPRP dataset. The complete dataset can be also found at https://github.com/steibelj/GWA_growth.

The dataset comes as `gpData` from the `synbreed` R package.

This example will demonstrate how to create pedigree based relationship matrix and adaptive window based LD from the dataset.

Finally, we'll create the data file for BATools.

### 0. Before you start
Make sure you have the following R package installed: `dplyr`,`MCMCglmm`, `Matrix`


### 1. Explore the data 

```{r}
load("MSUPRP_sample.RData")
class(MSUPRP_sample)
str(MSUPRP_sample)
```

### 2. Merge phenotype and covar to a dataframe
Commonly, using dataframe or datatable to manage all phenotypes are easier. So we'll merge the two together
```{r}
library(dplyr)
tmp<-data.frame(MSUPRP_sample$pheno)
tmp$id<-rownames(tmp)
colnames(tmp)=c("ph_24h"," temp_24h","driploss","id")
pheno<-left_join(tmp,MSUPRP_sample$covar,by="id")
head(pheno)
```


### 3. Check genotype matrix
```{r}
geno<-MSUPRP_sample$geno
dim(geno)
geno[1:5,1:10]
```

In BATools, the `id` will be matched with the `colname` of the genotype matrix, make sure you have genotype for all the individuals unless you're using single-step approach.

```{r}
genof0<-geno[which(as.numeric(rownames(geno))>6000),]
dim(genof0)
maf.f0<-colMeans(genof0,na.rm = T)/2
summary(maf.f0)
```


### 4. Create pedigree based additive genomic relationship matrix
```{r}
ped<-MSUPRP_sample$pedigree
head(ped)
```

```{r}
#source("tools.R")
library(pedigree)
colnames(ped)=c("id","father","mother","generation","sex")
ped0=Pedigree(ped,unknown = 0)
Ainv<-MCMCglmm::inverseA(pedigree=ped0[,1:3],nodes="ALL")$Ainv
A=Matrix::solve(Ainv,sparse=TRUE,tol=1e-16)
A[which(A<1e-6)]=0
A
```


### Building map with adaptive window based on LD
```{r}
map<-MSUPRP_sample$map
head(map)
```

Install `BALD` package from http://www.math-evry.cnrs.fr/logiciels/bald. On windows, `RTools` need to be pre-installed https://cran.r-project.org/bin/windows/Rtools/ and add your `RTools` path to enviroment variable. But linux or Mac are recommended since `BALD` is not available at CRAN. 

Installing these packages can be complicated, if you have any question, send me an email at chench57{at}msu.edu

```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("chopsticks")
biocLite("snpStats")
biocLite("ROC")
install.packages(c("LDheatmap","quadrupen", "ROC", "grplasso","snpStats"))
system("wget http://www.math-evry.cnrs.fr/_media/logiciels/bald_0.2.1.tar.gz")
system("R CMD INSTALL bald_0.2.1.tar.gz")
```
```{r}

```

